---
title: "Exploring Federal Opportunity Zones & Low-income Communities in the US"
output: 
  html_document:
    code_folding: hide
  bookdown::html_document2:
    css: styles.css
    fig_caption: yes
    theme: flatly
    toc: yes
    toc_depth: 1
bibliography: your-bib-file.bib
csl: computer.csl
link-citations: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(width = 50)

```

#INTRODUCTION
In 2017, the Trump Administration passed the Tax Cuts and Jobs Act [@usgov2017], authorizing the establishment of new "Opportunity Zones" (OZ) -- offering a new type of tax-incentive to target private equity investments into distressed communities. Of the 30,770 low-income community (LIC)[^LIC.note] census tracts eligible for designatoin, 8,700 were selected by state officials[^OZ.note], and approved by the US Treasury [@tres2018].  This program purports to offer an unprecedented level of opportunity for new investment aimed at community and economic development, with an estimated **$6.1 trillion** in capital elligible for this type of tax-exemption [@eig2018a]. Social impact funds, community development financial institutions, and state and local agencies across the country are now mobilizing to attract private funding, and to green-light shovel ready development projects [@article1]. However, with this new rush to reinvest and redevelopment are myriad concerns about how to manage new investment, how to prioritize community needs, and how to stave off negative effects associated with gentrification, displacement, and affordability. The Federal government has issued minimal guidance - both regarding how tracts were selected, and in terms of the types of projects that will be developed [@article2], and in the absence of and rules and regulations, many communities feel that the program may ultimately serve developers and investors, rather than the communities and people living in the OZ neighborhoods [@article3]. An initial assessment of designated tracts was conducted by the Economic Innovation Group [@eig2018b], but further analysis can provide a closer look at these communities, with some of that work to understand the selection process and potential already being conducted [@urban2018].

[^LIC.note]:Low-Income Community(LIC): LIC census tracts have an individual poverty rate of 20% or more, AND median family income up to 80% percent of the area median (area = state or metro).[@usgov2004]
[^OZ.note]:Opportunity Zone(OZ): Every state can designate up to 25% of its eligible census tracts. An exception permits 5% of non-LIC tracts may be eligible if contiguous with a LIC designated OZ tract (LIC-OZ), AND median family income does not exceed 125% of the median family income of the contiguous LIC-OZ tract.[@usgov2017]

#OVERIVEW
The goals of this project are to:

  1. Examine the **spatial distribution** of designated tracts
  2. Compare **community characteristics** of designated vs. non-designated (but eligible) tracts
  3. Assess **variability in 'level of need'**, based on multiple definitions of area disadvantage
  4. Evaluate the **potential for development** based on prior social impact investments

#DATA
##### Opportunity Zones
* _description:_ This national dataset includes census tract classifications of OZ eligibilty and designation.
* _source:_ US Treasury
* _access:_ [Opportunity Zones Resources](https://www.cdfifund.gov/Pages/Opportunity-Zones.aspx)
* _timeframe:_ 2018
* _geography:_ census tract
* _indicators:_
  `Eligible census tracts`
	`Approved designated census tracts`
		
#####  Community Characteristics 
* _description:_ These national databases are compiled annually for 5-year survey estimates, includes aggregate community characteristics and demographics.
* _source:_ American Community Survey Summary File (US Census)
* _link:_ [American Community Survey Summary File Data](https://www.census.gov/programs-surveys/acs/data/summary-file.html)
* _timeframe:_ annual 2010-2016 (5-year estimates)
* _geography:_ tract, country, state
* _indicators:_
	`demographics (age, sex, race, ethnicity, citizenship)`
	`socioeconomic (education, income)`
	`households (size, family type)`
	`housing (% urban, housing type, year built, vacancy)`
* spatial boundary files: TIGER/Line®[https://www.census.gov/geo/maps-data/data/tiger-line.html]

#####  New Markets Tax Credit (NMTC) Investments 
* _description:_ This national dataset includes all NMTC investment projects to present day, including date, location, project info (cost, type).
* _source:_ CDFI Fund
* _link:_ [CDFI Fund, NMTC Public Data Release](https://www.cdfifund.gov/news-events/news/Pages/news-detail.aspx?NewsID=270&Category=Press%20Releases)
* _timeframe:_ 2003-2015
* _geography:_ street address
* _indicators:_
  `Community Development Entity (CDE) Name` 
  `QALICB Type` 
  `Origination Year` 
  `Project QLICI Amount` 
  `Estimated Total Project Cost`

#METHODS
**Data processing** was conducted in R. That processing can be [viewed here](https://adamwilsonlabedu.github.io/geo503-2018-finalproject-dnkarp/data.html). Scripts and dataset can also be access through the projects [Github Repo](https://github.com/AdamWilsonLabEDU/geo503-2018-finalproject-dnkarp/tree/master/data). Note that ACS data accessed through `tidycensus` requires an API key from the Census, which can be [requested here](https://api.census.gov/data/key_signup.html). The R code includes a step where you would need to install that API key before being able to re-run those data pulls.
**R libraries** used in this analysis are listed in the codeblock below. These include tools for data wrangling, statistical analysis, spatial data management, analysis and visualization, and other graphics.
**Analytic steps** include importing and joining data from various geographic levels (tract, county, state), stratifying data by community type (low-income), grouping and summarizing new variables (designated tracts per county), plotting distributions, mapping geographic location of data, and statistically comparing subsets of the data (OZ vs. LIC). A secondary analysis identifies spatial neighbors and presents an additional computation of 'area disadvantage' (poverty ratio). Lastly, two logistic regression models are fitted to describes association between various input factors (community demographics, characteristics, and prior investment) and the output, or response, variable (OZ designation).

#ANALYSIS
##1. Libraries
Install and load libraries required for analysis.
```{r 1. load libs, echo=T, message=F, warning=F, cache=T, eval=T}

# uncomment the next line in order to install missing packages
# setup
# install.packages(c("devtools","tidyverse", "broom", "data.table", "purrr", "stringr", "skimr", "units", "sf", "maps", "viridis", "gridExtra", "ggthemes", "cowplot", "tableone", "ggpubr", "nngeo", "doMC", "readxl", "mapview","bookdown"))

library(devtools)
# install_github("hrbrmstr/albersusa")

# system
library(doMC)
library(bookdown)

# data wrangling libs
library(tidyverse)
library(broom)
library(data.table)
library(purrr)
library(stringr)
library(readxl)

# data analysis
library(skimr)
library(units)
library(tableone)

# spatial libs
library(sf)
library(maps)
library(albersusa)
library(nngeo)
library(mapview)

# graphics/viz libs
library(viridis)
library(gridExtra)
library(ggthemes)
library(cowplot)
library(ggpubr)
library(kableExtra)

```

##2. Import Data
Import cleaned data required for analysis. For more info, see steps in pre-processing [here](https://adamwilsonlabedu.github.io/geo503-2018-finalproject-dnkarp/data.html).
```{r 2 load data, message=F, warning=F, cache=T, eval=T}

# load OZ data frame (50 states + DC, territories exluded)
ozones.51 <- readRDS("data/ozones_51.rds")

# load US census tracts sf data frame
tracts.sf <- readRDS("data/tracts_sf.rds")

# load US counties sf data frame
counties.sf <- readRDS("data/counties_sf.rds")

# load US states sf data frame
states.sf <- readRDS("data/states_sf.rds")

# load nmtc tract data
nmtc.tract <- readRDS("data/nmtc_tract.rds")

# load tract census attributes
us.trcts.acs.rt16 <- readRDS("data/us_trcts_acs_rt16.rds")

# load county census attributes
us.cnty.acs.rt16 <- readRDS("data/us_cnty_acs_rt16.rds")

# load state census attributes
us.st.acs.rt16 <- readRDS("data/us_st_acs_rt16.rds")

# load rural urban continuum code county data
ruralurbancodes2013 <- readRDS("data/ruralurbancodes2013.rds")

```

##3. Join attributes data to spatial features
Use `merge` function to join Opportunity Zone data, NMTC data, ACS, RUCC data to the census tracts `sf` object.
```{r 3 merge data, message=F, warning=F, cache=T, eval=T}

# merge (left join) 'Opportunity Zones' to sf tracts dataframe ; merge NMTC data and ACS attributes to Ozone tracts sf df
sfdf <- merge(tracts.sf, ozones.51[,-c("state","state_code")], by.x = "GEOID", by.y = "GEOID10", all.x) %>% # merge OZ designations to lower 48 tracts sf data frame
  select(-c(variable,estimate,moe)) %>% 
  merge(nmtc.tract, by="GEOID", all.x=T) %>% # merge NMTC data to tracts
  merge(us.trcts.acs.rt16,by="GEOID", all.x=T) %>% # merge ACS varaibles (cleaned 2016 rates)
  mutate(CountyFIPS=substr(GEOID,1,5)) %>% 
  merge(us.cnty.acs.rt16[,c("GEOID","PovRt","MdnInc")], by.x="CountyFIPS",by.y="GEOID", all.x=T) %>% # merge country level variables to tracts
  rename(PovRt.trct = PovRt.x) %>% 
  rename(MdnInc.trct = MdnInc.x) %>% 
  rename(PovRt.cnty = PovRt.y) %>% 
  rename(MdnInc.cnty = MdnInc.y) %>% 
  mutate(StateFIPS=substr(GEOID,1,2)) %>% 
  merge(us.st.acs.rt16[,c("GEOID","PovRt","MdnInc")], by.x="StateFIPS",by.y="GEOID", all.x=T) %>% # merge state level variables to tracts
  rename(PovRt.st = PovRt) %>% 
  rename(MdnInc.st = MdnInc) %>% 
  mutate(PovRatio.cnty = PovRt.trct/PovRt.cnty) %>% 
  mutate(PovRatio.st = PovRt.trct/PovRt.st) %>% 
  mutate(MdnIncRatio.cnty = MdnInc.trct/MdnInc.cnty) %>% 
  mutate(MdnIncRatio.st = MdnInc.trct/MdnInc.st) %>% 
  mutate(OZ = factor(ifelse(QOZ=="DESIGNATEDQOZ", "OZone", ifelse(TYPE=="LIC", "Non-OZ LIC", "Non-LIC"))))

# join RUCC data, and recode 'urban' classification
sfdf <- left_join(sfdf,ruralurbancodes2013[,c(1,5:6)],by=c("CountyFIPS"="FIPS")) 
sfdf <- sfdf %>% mutate(urban = (RUCC_2013 %in% c(1,2,3))) 

```

##4. Descriptive analysis
###4a. National and State Distribution
Generate tables to show the distribution of census tract eligibility: non-eligibile, eligible (low-income), and eligible (contiguous) by OZ designation. Use `kable` to format tables.
```{r 4a1 eligibilty table 1, message=F, warning=F, cache=T, eval=T, results='hide'}

# two-way table of eligibility and designation
table(sfdf$QOZ,sfdf$TYPE) %>% #counts
  kable() %>%
  kable_styling()

```

```{r 4a1 eligibilty table 2, message=F, warning=F, cache=T, eval=T, results='hide'}

# two-way table of eligibility and designation
prop.table(table(sfdf$QOZ,sfdf$TYPE)) %>% #proportions
  kable() %>%
  kable_styling()

```

**Table 1** shows counts by eligibilty x OZ designation.
```{r 4a1 display table 1, message=F, warning=F, cache=T, eval=T, ref.label='4a1 eligibilty table 1'}
```

**Table 2** shows proportions by eligibility x OZ designation.
```{r 4a1 display table 2, message=F, warning=F, cache=T, eval=T, ref.label='4a1 eligibilty table 2'}
```

Generate a bar chart showing the distribution of designated OZ tracts, non-OZ low-income tracts, and non-eligible tracts. Sort by OZ proportion per state, and use `ggplot` to create **Figure 1**.
```{r 4a2 tract type by state bar chart, message=F, warning=F, cache=T, eval=T}

# bar chart: state level distribution of tract designation/eligibility (sorted by % OZ)
sfdf %>% 
  st_set_geometry(NULL) %>% 
  group_by(state) %>% 
  summarise(n = n(), OZone = sum(OZ == "OZone")/n,LIC = sum(OZ == "Non-OZ LIC")/n,NonLIC = sum(OZ == "Non-LIC")/n) %>% 
  mutate(state = factor(state, levels = state[order(OZone)])) %>% 
  gather(key = "OZdesignation", value = "n.tracts", OZone, LIC, NonLIC) %>% 
  ggplot(aes(x = state, y = n.tracts, fill = factor(OZdesignation, levels = c("NonLIC","LIC","OZone")))) + 
  theme_bw() + geom_bar(stat = "identity") + 
  ylab("% of state tracts") + 
  labs(fill = "Census Tract\nEligibility/Designation") + 
  scale_fill_manual(values=c("grey","blue","dark blue")) +
  theme(text = element_text(size=10), axis.text.x = element_text(angle=90, hjust=1))

```   

Generate maps of tracts colored by eligibile type, using `ggplot` and `geom_sf`.  Define a function `mkmaps` that takes parameters to make tract, county, or state level maps, and allows subset by state ID(s) (e.g. "AL").
```{r 4a3 generate maps, message=F, warning=F, cache=T, eval=T}

# define map layout
noaxis <- theme(
  axis.title.x=element_blank(),
  axis.text.x=element_blank(),
  axis.ticks.x=element_blank(),
  axis.title.y=element_blank(),
  axis.text.y=element_blank(),
  axis.ticks.y=element_blank(),
  panel.background = element_blank(),
  legend.title = element_text(size=8),
  legend.text = element_text(size=8)
  )

# definte map function
mkmaps <- function(x,y,z,w){
x %>%
  filter(state %in% z) %>% 
  group_by(eval(parse(text=w))) %>% 
  summarise(value = first(eval(parse(text=y)))) %>% 
  ggplot(aes(fill = value)) +
  geom_sf(color = NA, lwd=.5) +
  scale_fill_manual(values=c("#56B4E9", "#E69F00", "#999999")) +
  labs(fill=y) +
  ggtitle("Distribution Tracts by Eligibility") +
  geom_sf(data = st_geometry(states.sf[states.sf$state %in% z,]), fill = "transparent", color = "white", lwd=.5) +
  geom_sf(data = st_geometry(x[(x$state %in% z) & x$OZ=="OZone",]), fill = "transparent", color = "red", lwd=.75) +
  noaxis %>%
    return()
}

mkmaps2 <- function(x,y,z,w){
x %>%
  filter(state %in% z) %>% 
  group_by(eval(parse(text=w))) %>% 
  summarise(value = first(eval(parse(text=y)))) %>% 
  ggplot(aes(fill = value)) +
  geom_sf(color = NA, lwd=.5) +
  labs(fill=y) +
  ggtitle("Poverty Rate and Designated OZone Tracts") +
  geom_sf(data = st_geometry(states.sf[states.sf$state %in% z,]), fill = "transparent", color = "white", lwd=.5) +
  geom_sf(data = st_geometry(x[(x$state %in% z) & x$OZ=="OZone",]), fill = "transparent", color = "red", lwd=.75) +
  noaxis %>%
    return()
}

# set-up additional variables for mapping
s <- sfdf %>% 
  mutate(area = set_units(st_area(.),mi^2)) %>% 
  mutate(QOZ = as.factor(QOZ)) %>%  #set QOZ as factor 
  mutate(popdens = as.double(Pop/area)) %>% 
  mutate(prjdens = as.double(n.prj/area)) %>% 
  mutate(invdens = as.double(ttl.amt.prj/area))

```

Create map of tracts by eligibility type for continental US. _(DO NOT RUN, CURRENTLY SET TO FALSE)_
```{r 4a3 national map, message=F, warning=F, cache=T, eval=F}

# create map object for enture US
s %>% st_transform(3857) %>% mkmaps('TYPE',unique(s$state),"GEOID")

```

Create map of tracts by eligibility type for PA, NY and NJ. **Figure 2**
```{r 4a3 state map, message=F, warning=F, cache=T, eval=T}

# create map object for enture US
s %>% st_transform(3857) %>% mkmaps('TYPE',c("PA","NY","NJ"),"GEOID")

```

Create interactive map of tracts by OZ designation type for PA, NY and NJ, using `leaflet` and `mapview`.
```{r 4a3 interactive map, message=F, warning=F, cache=T, eval=T}

sfdf[sfdf$state %in% c("PA","NY","NJ"),] %>% mapview(zcol="QOZ",legend=T)

```

Review distribution of OZ designations per county by state, include urban/non-urban classification.  Urban defined by RUCC codes 1,2,3: counties in large metro areas. **Table 3**
```{r 4a4 tracts per county state summary, message=F, warning=F, cache=T, eval=T, size="tiny"}

# summarize state distribution for tracts per country
sfdf %>% 
  st_set_geometry(NULL) %>% 
  group_by(CountyFIPS) %>% 
  summarise(StateFIPS = first(StateFIPS), urban = first(urban), trcts.per.cnty.n = n(), oz.per.cnty.n = sum(OZ == "OZone"), oz.per.cnty.pct = oz.per.cnty.n/trcts.per.cnty.n) %>% 
  group_by(StateFIPS) %>% 
  summarise(urb.oz.per.st.n = sum(oz.per.cnty.n * urban), oz.per.st.n = sum(oz.per.cnty.n), urb.oz.per.st.pct = urb.oz.per.st.n/oz.per.st.n) %>% 
  select(oz.per.st.n, urb.oz.per.st.pct) %>% 
  skim()

```

Plot percentage of OZ designations per county by state in a box plot, stratified by Urban/non-urban. **Figure 3**
```{r 4a5 tracts per county state box plots, message=F, warning=F, cache=T, eval=T, fig.width=20}

# create box plot of state distribution for tracts per country
x <- sfdf %>% 
  st_set_geometry(NULL) %>% 
  group_by(CountyFIPS) %>% 
  summarise(state = first(state), urban = first(urban), trcts.per.cnty.n = n(), oz.per.cnty.n = sum(OZ == "OZone"), oz.per.cnty.pct = oz.per.cnty.n/trcts.per.cnty.n)  
  ordered_state = with(x, reorder(state, oz.per.cnty.pct, median))

  ggplot(x, aes(ordered_state, oz.per.cnty.pct, fill=factor(urban))) +
  geom_boxplot() +
  geom_boxplot(outlier.alpha = 0.1) +
  ylab("% County OZ tracts") + 
  labs(fill = "Urban/Rural") +
  theme(text = element_text(size=8), axis.text.x = element_text(angle=90, hjust=1)) + 
  scale_fill_manual(values=c("#E69F00", "#56B4E9"))

```   

###4b. Compare Designated and Eligible Tracts pt1
Subset data to select designated tracts and non-designated, but eligible tracts. Stratify and compare the two groups to describe area characteristics, prior investment activity, income/poverty level indicators that are used in eligibility criteria. This step aims to explore characteristics that differentiate designated OZ tracts to understand variables that may have factored into selection.
```{r 4b1 Compare OZ and Non-OZ-LIC pt1,  message=F, warning=F, cache=T, eval=T}

# select dataset: remove geometry (increase process speed), filter by LIC only, add area/density variables
d <- sfdf %>% 
  mutate(area = set_units(st_area(.),mi^2)) %>% 
  st_set_geometry(NULL) %>% 
  mutate(QOZ = as.factor(QOZ)) %>%  #set QOZ as factor 
  filter(QOZ=="DESIGNATEDQOZ" | TYPE == "LIC") %>% #restrict to LIC (low-income communities == QOZ eligible)
  mutate(OZ = factor(OZ)) %>% 
  mutate(popdens = as.double(Pop/area)) %>% 
  mutate(prjdens = as.double(n.prj/area)) %>% 
  mutate(invdens = as.double(ttl.amt.prj/area))

```

Use `CreateTableOne` to generate descriptive tables, compute univariate test statistics. These tables compare all variables between OZ and Non-OZ-LIC (ChiSquared tests are used for categorical variables; One-way ANOVA tests are used for continuous variables). Tables include national totals.

**Table 4** shows demographic (ACS) characteristics stratified by tract status
```{r 4b2 Compare OZ and Non-OZ-LIC pt1,  message=F, warning=F, cache=T, eval=T, size="tiny"}

# demographic mix
tableOne1 <- CreateTableOne(strata = "OZ", vars = c("Pop","popdens","PovRt.trct","MdnInc.trct","UnemplRt","VacRt","RentRt","GrEduRt","ColEduRt","PctWhite"), data = d)
tableOne1

```

**Table 5** shows prior investment (NMTC) characteristics stratified by tract status
```{r 4b3 Compare OZ and Non-OZ-LIC pt1, message=F, warning=F, cache=T, eval=T, size="tiny"}

# prior investment
tableOne2 <- CreateTableOne(strata = "OZ", vars = c("metro","n.cde","n.prj","avg.prj.yr","ttl.amt.prj","avg.amt.prj","avg.amt.yr","CDE","NRE","RE","SPE"), data = d)
tableOne2

```

**Table 6** shows 'distressed area' criterion (income and poverty characteristics) stratified by tract status
```{r 4b4 Compare OZ and Non-OZ-LIC pt1, message=F, warning=F, cache=T, eval=T, size="tiny"}

# distressed area criterion
tableOne3 <- CreateTableOne(strata = "OZ", vars = c("PovRatio.cnty","PovRatio.st","MdnIncRatio.cnty","MdnIncRatio.st","prjdens","invdens"), data = d)
tableOne3

```

**Figure 4** shows distributions of poverty levels stratified by tract status
```{r 4b5 Compare OZ and Non-OZ-LIC pt1, warning=F, message=F, cache=T, eval=T}

# compare distribution of Poverty Rate between OZ and Non-OZ LIC
d %>% ggplot(aes(PovRt.trct, color = OZ)) + 
  scale_color_manual(values=c("#E69F00", "#56B4E9"), 
                       name="Census Tract\nDesignation",
                       breaks=c("OZone","Non-OZ LIC"),
                       labels=c("Designated OZone", "Non-OZone LIC")) +
  geom_density(alpha = 0.4) + 
  geom_vline(data=data.frame(d[!is.na(d$PovRt.trct),] %>% 
                               group_by(OZ) %>% 
                               summarise(grp.mean=mean(PovRt.trct))), 
    aes(xintercept=grp.mean, color=OZ), 
    linetype="dashed") +
    xlab("2016 Poverty Rate") 
  
```                     

###4c. Compare Designated and Eligible Tracts pt2
Explore select variables, comparing distributions between designated and non-designated tracts. Use `ggboxplot` to generate plots. **Figure 5**
```{r 4c Compare OZ and Non-OZ-LIC pt2,  message=F, warning=F, cache=T, eval=T}

# review distribution of variables
# d.skim <- skim(d[23:39])

# plot continuous variables x bin/cat vars >> box plots (statistically significant vars X OZ class)

p1 <- ggboxplot(d, x = "OZ", y = "PovRt.trct",
                 color = "OZ", palette = "jco")
p2 <- ggboxplot(d, x = "OZ", y = "UnemplRt",
                 color = "OZ", palette = "jco")
p3 <- ggboxplot(d, x = "OZ", y = "VacRt",
                 color = "OZ", palette = "jco")
p4 <- ggboxplot(d, x = "OZ", y = "GrEduRt",
                 color = "OZ", palette = "jco")
p5 <- ggboxplot(d, x = "OZ", y = "ColEduRt",
                 color = "OZ", palette = "jco")
p6 <- ggboxplot(d, x = "OZ", y = "PctWhite",
                 color = "OZ", palette = "jco")

ggarrange(p1,p2,p3,p4,p5,p6 + rremove("x.text"), 
          labels = c("A", "B", "C", "D", "E", "F"),
          ncol = 3, nrow = 2)

```                     

##5. Map data
Generate tract, county, and state level poverty rate maps for Alabama and Georgia. Add layer showing locations (in orange).  Using `mkmaps` function defined above. **Figures 5,6,7**
```{r 5 generate maps, message=F, warning=F, cache=T, eval=T}

# Map poverty data (functions defined above)
tract.povrt.map <- s %>% mkmaps2('PovRt.trct',c("AL","GA"),'GEOID')
county.povrt.map <- s %>% mkmaps2('PovRt.cnty',c("AL","GA"),'CountyFIPS')
state.povrt.map <- s %>% mkmaps2('PovRt.st',c("AL","GA"),'StateFIPS')

tract.povrt.map
county.povrt.map
state.povrt.map

```

##6.Calculating Area Poverty
```{r sample table, message=F, warning=F, cache=T, eval=T, results='hide'}
data.frame(GEOID=c(1,1,1,1,2,2,2,2),neighbor=c(3,6,5,8,2,4,5,6)) %>% 
  kable() %>%
  kable_styling()

```

Define function to identify nearest neighbors. Two functions are written: `neighb.do` for serial and `neighb.dopar` for parallel processing.  To use the parallel option, `regiterDoMC` is used to sepecify the number of cores to employ. Input parameters are required to specify the dataset, number of rows to analyze, number of nearest neighbors, and maximum search distance. The functions return a `data.frame` object in a long format, e.g. 

```{r show table, , message=F, warning=F, cache=T, eval=T, ref.label='sample table'}
```

```{r 6 define neighbor average function, message=F, warning=F, cache=T, eval=T}

# register 4 cores for parellel processing
registerDoMC(cores=4)

# define average neighbor poverty functions

# serial
neighb.do <- function(dt,n,knn,dis){
  neighb <-foreach(i=dt[1:n,]$GEOID, .combine='rbind') %do%
  {
    r <- dt %>% 
      filter(GEOID == i) %>% 
      st_nn(dt, sparse = TRUE, k = knn, maxdist = dist, returnDist = FALSE, progress = TRUE) 
    dim(r) <- c(1,1)
    rownames(r) <- i
    data.frame(rowname=names(r[,1][i]),neighbor=r[,1][[i]])
  }
  return(neighb)
}

# parallel
neighb.dopar <- function(dt,n,knn,dist){
  neighb <-foreach(i=dt[1:n,]$GEOID, .combine='rbind') %dopar%
  {
    r <- dt %>% 
      filter(GEOID == i) %>% 
      st_nn(dt, sparse = TRUE, k = knn, maxdist = dist, returnDist = FALSE, progress = TRUE) 
    dim(r) <- c(1,1)
    rownames(r) <- i
    data.frame(rowname=names(r[,1][i]),neighbor=r[,1][[i]])
  }
  return(neighb)
}

```

Compute average area poverty using the neighbors matrix. Alabama and Georgia are run as a demonstration, search for 10 nearest neighbors, within a 5km maximum distance. The neighbor average is computed, and then a ratio for tract to neighbors is calculated.
```{r 6 compute area poverty, message=F, warning=F, cache=T, eval=T}

# subset data to Alabama and Georgia
d <- sfdf %>% 
  filter(state %in% c("AL","GA")) %>% 
  select(GEOID,PovRt.trct) %>% 
  st_transform(3857)

# get neighbors (AL + GA, all tracts, 10 nearest neighbors, 5km max)
neighb <- neighb.dopar(d,nrow(d),10,5000)

# compute average poverty for neighbors
df <- rownames_to_column(d) %>% 
  st_set_geometry(NULL) 

nf <- neighb %>% 
  mutate(rowname = as.character(rowname), neighbor = as.character(neighbor))

dx <- left_join(nf,df,by=c("neighbor"="rowname")) %>% 
  mutate(PovRt.nb = ifelse(rowname != GEOID,PovRt.trct,NA)) %>%
  mutate(PovRt.trct = ifelse(rowname == GEOID,NA,PovRt.trct)) %>%
  mutate(neighbor = ifelse(rowname == GEOID,NA,neighbor)) %>% 
  select(-GEOID) %>% 
  rename(GEOID = rowname) %>% 
  group_by(GEOID) %>% 
  summarise(PovRt.trct = max(PovRt.trct, na.rm = T), neighb.cnt = n(), PovRt.area = mean(PovRt.nb, na.rm = T)) %>% 
  mutate(pov.area.ratio = PovRt.trct/PovRt.area)

sfdf.ALGA <- merge(sfdf, dx, by = "GEOID", all.y = T) %>% 
  select(-PovRt.trct.y) %>% 
  rename(PovRt.trct = PovRt.trct.x)

```

**Figure 8** maps the the poverty rate per tract, the _local_ area poverty rate (average of neighbors) and the poverty rate ratio between each tract and it's local neighbors.
```{r 6 plot area poverty, message=F, warning=F, cache=T, eval=T}

# plot data
sfdf.ALGA %>% 
  select(PovRt.trct,PovRt.area,pov.area.ratio) %>% 
  plot()

```

##7.Regression Analysis
Fit 2 regression models that estimate probability of eligible tracts being designated as OZones.  Input variables include area demographic characteristics and prior investments (see **section 4** above).
```{r 7 regression analysis, message=F, warning=F, cache=T, eval=T}

# fit logistic regression for area demographics
d2 <- sfdf %>% 
  filter(OZ!="Non-LIC") 

summary((glm(OZ~Pop + PovRt.trct + MdnInc.trct + UnemplRt + VacRt + RentRt + ColEduRt + GrEduRt + PctWhite, data=d2, family=binomial())))

# fit logistic regression for area investments
summary((glm(OZ~n.cde + n.prj + avg.prj.yr + ttl.amt.prj + avg.amt.prj + avg.amt.yr, data=d2, family=binomial())))

# report odds ratios for poverty rate
exp(coef(glm(OZ~PovRt.trct, data=d2, family=binomial())))

```

#SUMMARY OF FINDINGS

Of the 7,773 tracts designated as Qualified Opportunity Zones (OZ), 97% were classified as Low-Income (LIC), which represents about 10% of all tracts in the counrty. Additionally, 195 were designated based on criteria for contiguous tracts, and another 17 granted some other status for exception (see **Table 1** and **Table 2**). Note that these totals currently exclude Alaska, Arizona, Puerto Rico, and other outlying territories (adds approximately 1,000 additional designated OZ tracts). The distribution for the lower 48 states, plus Washington DC, ranges for total LIC tracts from around 25% to as high as 60% (in Mississippi), while states designated between 10 and 20% of total tracts. The proportion of eligible tracts that were designated ranges from about 25% (IA) to 75% (WY) (see **Figure 1**). The map presented in **Figure 2** allows a visual examination of the geographic distribution of selected and eligible tracts across NY, PA, and NJ. While much of the rural areas of western PA, upstate NY, and southern NJ were eligible, we can see clustering in NYC and Philadelphia areas, as well as Pittsbugh, Buffalo and various areas into the Hudson Valley and along the NY Thruway. For closer inspection, the interactive map can be used to zoom into these denser urban areas where census tracts are much smaller. 

To further inspect this question of urban and rural selection, **Table 3** shows state distributions of number of OZones (median=126, range=24-879) and proportion of OZones in urban counties (median=71%, range=16-100%). **Figure 2** helps visualize these patterns with a box plot showing urban and non-urban (includes lower population towns as well as rural). This figure plots each state's distribution of the proportion of each county's tracts designated as an OZone, and stratified by urban status of the county. States are order left to right based on the median county OZone percentage (for all counties combined) -- therefore, GA had the lowest median OZ tracts per county, whereas CO had the highest.  This is interesting considering that Coloardos urban counties had a relevatively low average and maximum percentage, but not surprising given that urban areas are denser and have more tracts in total.  So, rural county OZ designations are likely to cover more territory and proportionally more of the region, while urban designations are more likely to be sparse or perhaps more concentrated neighborhods, or perhaps industrial areas. 

To explore the characteristics of which areas were designated as OZones, it is useful to compare that population of tracts with those that were eligiible but not selected. presents some descriptive statistics about area demographics (**Table 4**), prior investment (**Table 5**), and area income and poverty (**Table 6**). Overall, OZones tend to be lower population density, higher poverty, lower median family income, higher unemployment, higher rate of vacancy, higher rate of renters, less educated, and less white. All but population showed statistically significant _(p<0.001)_ difference between the two groups. The investment indicators included variables related to the NMTC program.  Aggregate values were totaled from ~10 years of projects by location.  OZones tend to have greater numbers of Community Development Entities (CDE)[^CDE], more New Market Tax Credit NMTC projects[^NMTC], more projects per year, higher dollar investment, higher dollar investment per project and per year.  All of these variables were statistically signficant at _p<0.001_, except amount per project was signficant at _p<0.01_. **Figure 3** and **Figure 4** plot a selection of these variables, to aid in visualizing differences between these two populations.

**Table 6** considers area income and poverty values, which are used as indicators for Low-Income Community (LIC) status. The ratio of tract-level poverty to state and county poverty, and the ratio of tract-level median family income to state and county median family income were computed.  OZone tracts had higher poverty ratios for both state and county, and lower median income for state and county.  All were statistically significant at _p<0.001_.  The average poverty rate for OZone tracts was nearly 2 times higher than the state rate, and 1.8 time higher than the county, compared with 1.6 and 1.5, for Non-OZone tracts, respectively. These ratios are explored further, by calculating a more localized, non-jurisdictional approach to area disadvantage (see **section 6**).

Using Alabama and Georgia for demonstration **Figures 5-7** map the poverty rate for tract, county, and state, respectively, including a layer showing the distribution of designated OZones tracts. In both states there are concentrations of OZones in Birmingham and Atlanta, with a visible number circling the outskirts of both metro areas.  More notable, both states have a signficant spread across the southern regions where we can see county level poverty increase. 

The code developed to compute neighbor averages can be tested further. For example, the number of neighbors and the optimal bandwidth can be explored with a sensitivity analysis to identify where these results overlap with county and state metrics. It is important to note that metro area rates can be obtained from the census (although those data are not considered in this analysis at this time).  However, these additional, hyperlocal, or more scale flexible definitions of area disadvantage may be more valuable in smaller towns, rural communities, and especially inter-state regions. The maps shown in **Figure 8** demonstrate an initial result of this process.

Another area of this analysis to develop further is the regression modeling.  Using the area indicators presented in **section 4**, logistic regression models were fit to estimate outcome of whether an eligible tract was designated as an OZone. Similar to the results in **section 4**, probability of OZone designated was associated with increases in population, poverty, unemployment, vacancy, rent, and less education. An increase in percent white and median income were also found to improve likelihood of OZ designation, which could perhaps be a result of non-LIC exemptions (contiguous status), or some other issues in the model (colinearity?) which would require further tuning and calibration. Also associated were higher numbers of Community Development Entities and higher numbers of New Market Tax Credit (NMTC) projects, per annum.

[^CDE]: A CDE is a domestic corporation or partnership that is an intermediary vehicle for the provision of loans, investments, or financial counseling in low-income communities.
[^NMTC]: NMTC investments are made into various kinds of projects including real estate and non-real estate (other business financing).

#CONCLUSION
This analysis provides a starting point to understanding this new policy and investment landscape as it evolves in the coming years.  OZ tracts were chosen by state officials, with minimal required selection critieria. Without knowing the exact critieria used in each state, this analysis serves to describe the ultimate outcome of those selection processes, to present an overview of these investment-eligible communities, and to consider some of the variables relevant to evaluating the potential impact of the OZ program. As development projects begin to breakground much research should be done to understand how local and regional governments mobilize to leverage this program as an economic development tool.  A question still remains whether the federal government will collect any data on these projects, as it did for the NMTC program.  The lack of project level data will pose major limitations to such analysis, making ancillary and local level data sources vital. Future work may look to classify and subset OZ tracts based on previous patterns of upward or downward trajectory in terms of recent demographic and economic change.
  
#REFERENCES


